Using numerical modelling to investigate temperature and wave exposure as drivers of macroalgal characteristics

### 2019-08-04 morphology_paper.Rmd
methods section updated and added to

Site-specific summary statistics were calculated for each of the environmental variables considered in this study. The summary statistics calculated for temperature, waves and wind variables were minimum, maximum, mean, range and standard deviation for annual, February (Austral summer) and August (Austral winter) timescales, respectively. In addition, annual and seasonal median wave and wind direction were also calculated; all summary data and their respective abbreviation are presented in Table 3. The median was calculated for wind and wave direction, as issues arise when calculating the mean and standard deviation for compass metrics. Dot and whisker plots were used to present the summary statistics for both temperature and wave variables. Summary statistics for wind was not plotted and instead are discussed, as the data was course relative to the other environmental variables considered in this study.

*insert paragraph about morphology*
Characterising and comparison of morphological characteristics between deep and shallow *E. maxima* was achieved by notched boxplots. Distance-based redundancy analysis was performed using the *rda* function in the *vegan* software package (Oksanen et al. 2016) in R (R Core Team, 2019). The role of temperature, waves and wind as drivers of kelp morphology was investigated. The abiotic data covered all study sites for both species of kelp. Both morphology and abiotic data (temperature, waves and wind) were standardised using the *decostand* function in the *vegan* software package; morphology data were used as response variables and the abiotic data as explanatory variables. To determine the explanatory variables that best describe patterns in the response data, a full RDA was performed using a complete set of explanatory variables. Forward selection was then used to reduce the number of explanatory variables as well as prevent the inflation of overall type I error. To further improve the model, pairwise coefficients and Variance Inflation Factor (VIF) were calculated to identify variables with high multicollinearity. The computation of the parsimonious RDA was followed by permutation tests of the adjusted *R*2 to assess significance of constraints.

>>>>

Summary statistics were calculated for each site and regions for each of the environmental variables considered in this study. The summary statistics calculated for temperature, waves and wind variables were minimum, maximum, mean, range and standard deviation for annual timescales. In addition, annual median wave and wind direction were also calculated; all summary data and their respective abbreviation are presented in Table 3. The median was calculated for wind and wave direction, as issues arise when calculating the mean and standard deviation for compass metrics. Dot and whisker plots were used to present the summary statistics for both temperature and wave variables. Summary statistics for wind was not plotted and instead are discussed, as the data was course relative to the other environmental variables considered in this study.

Characterisation of morphological characteristics in relation to temperature and wave exposure non-parametric statistical methods were used. Morphological characteristics were categorised according to temperature regime and wave exposure category for each each species and population of kelp in this study. Significant differences between temperatre regime and wave exposure categories were tested with wilcox tests. Notched boxplots allowed comparison between individual categories of wave exposure. In order to investigate specific drivers of morphological characteristics a distance-based redundancy analysis was performed using the *rda* function in the *vegan* software package (Oksanen et al. 2016) in R (R Core Team, 2019). The abiotic data covered all study sites for both species of kelp. Both morphology and abiotic data (temperature, waves and wind) were standardised using the *decostand* function in the *vegan* software package; morphology data were used as response variables and the abiotic data as explanatory variables. To determine the explanatory variables that best describe patterns in the response data, a full RDA was performed using a complete set of explanatory variables. Forward selection was then used to reduce the number of explanatory variables as well as prevent the inflation of overall type I error. To further improve the model, pairwise coefficients and Variance Inflation Factor (VIF) were calculated to identify variables with high multicollinearity. The computation of the parsimonious RDA was followed by permutation tests of the adjusted *R*2 to assess significance of constraints.
